Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-28 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Madrzy_Thesis: update references

Now that the characteristics of the fires have been measured, the next step is to measure the thermal impact that the fires have on a target.  In these experiments the target is a non-combustible wall. The non-combustible material was chosen to eliminate any contribution from the wall to the fire.  The fires were initially placed at predetermined, fixed positions away from the wall.  In the remainder of the report the term burner will be considered synonymous with the pan for the liquid fuel or the block of solid fuel. Experiments were conducted with the closest edge of the burner located in the following positions from the wall: two burner widths (0.610~m), one burner width (0.305~m), one half burner width (0.150~m), and the touching the wall.

@@

A free standing steel frame wall with a 12.7~mm thick, 2.44~m high by 1.22~m wide, calcium silicate face was instrumented to measure the thermal exposure of the wall from the fire in terms of temperature and heat flux.  Two vertical arrays of water cooled, Schmidt-Boelter total heat flux gauges, as previously described in the instrumentation section above, were installed at 20 cm intervals from 0.2~m to 1.2~m above the burner.  One vertical array was centered on the burner and the other vertical array was centered on the right edge of the burner. The arrangement is shown in Fig.~\ref{Instrumented_NC_Wall}. Bare-bead, Chromel-Alumel (type K) thermocouples, with a 0.5~mm nominal diameter were installed next to the center of each heat flux sensors and continued, at 20~cm intervals, up to a height of 2.2~m above the surface of the burner. Photographs of the front and rear of the instrumented wall are shown in Figures~\ref{Instrumented_Wall_Front_photo} and \ref{Instrumented_Wall_Rear_photo} The thermocouples where installed in small holes from the back of the wall of the wall.  The bead of the thermocouple was flush with the surface on the calcium silcate board and was in contact with the board, as shown in Figure~\ref{Instrumented_Wall_Close_up_TC_HF_photo} For the center array,a second thermocouple was installed on the backside of the calcium silicate board, approximately 10~mm below the TC positioned on the front surface of the wall. The photograph in Figure~\ref{Instrumented_Wall_Detail_Rear_TC_photo} shows a rear thermocouple with the bead in contact with the rear surface of the wall and the relative position to the thermocouple on the front surface of the wall. In a manner similar to the source fire characterization experiments, the instrumented wall experiments were conducted under the NIST Large Fire Research Laboratory’s 3~m by 3~m oxygen depletion calorimeter.

@@

\section{Characterization of Gypsum Board}

@@

@@

@@

@@

\bibliography{../References/Madrzykowski_Thesis}
\end{document}

>>>>

Now that the characteristics of the fires had been measured, the next step was to measure the thermal impact that the fires had on a vertical target surface.  In these experiments the target is a non-combustible wall. The non-combustible material was chosen to minimize any contribution from the wall to the fire.  The fires were placed at predetermined, fixed positions away from the wall.  In the remainder of the report the term burner will be considered synonymous with the pan for the liquid fuel or the block of solid fuel. Experiments were conducted with the closest edge of the burner located in the following positions from the wall: two burner widths (0.610~m), one burner width (0.305~m), one half burner width (0.150~m), and then touching the wall (0.0~m).

@@

A free standing steel frame wall with a 12.7~mm thick, 2.44~m high by 1.22~m wide, calcium silicate board face was instrumented to measure the thermal exposure of the wall from the fire in terms of temperature and heat flux. Marinite Type I is the commercial name of the non-combustible board used in these experiments.  It is normally used as a fire-resistant thermal insulation that also has structural strength. 90~\% of the board is composed of calcium silicate and calcium meta-silicate.  The remaining 10~\% of the board is composed of a combination of organic fibers, fiber glass filament and crystalline silica (quartz) ref BNZ MSDS.  Material properties that impact heat transfer were selected from the manufacturer's literature and are presented in the table below ref BNZ Marinite I.   
\begin{table}
	\centering
\begin{tabular}{|c|c|c|c|c|}
	\hline Density (kg/m$^3&)  &Thermal Conductivity   & Specific Heat  &  &  \\ 
	\hline   & Temperature ($^\circ$C) & (W/mK)  & Temperature ($^\circ$C)  & (kJ/kg $^\circ$C) \\ 
	\hline 737  & 24 & 0.13 & 93 & 0.28 \\ 
	\hline  & 205 & 0.12  & 205 & 0.30 \\ 
	\hline  & 316 & 0.11 & 316 & 0.32 \\ 
	\hline  & 425 & 0.12 & 425 & 0.34 \\ 
	\hline  & 538 & 0.12 &  &  \\ 
	\hline 
\end{tabular} 
\caption[Thermal properties for Marinite I]{Thermal properties for Marinite I, as reported by the manufacturer BNZ Marinite I.}
\label{tab:Marinite_Thermal_Properties}
\end{table}

Two vertical arrays of water cooled, Schmidt-Boelter total heat flux gauges, as previously described in the instrumentation section above, were installed at 20 cm intervals from 0.2~m to 1.2~m above the burner.  One vertical array was centered on the burner and the other vertical array was centered on the right edge of the burner. The arrangement is shown in Fig.~\ref{Instrumented_NC_Wall}. Bare-bead, Chromel-Alumel (type K) thermocouples, with a 0.5~mm nominal diameter were installed next to the center of each heat flux sensors and continued, at 20~cm intervals, up to a height of 2.2~m above the surface of the burner. Photographs of the front and rear of the instrumented wall are shown in Figures~\ref{Instrumented_Wall_Front_photo} and \ref{Instrumented_Wall_Rear_photo} The thermocouples where installed in small holes from the back of the wall of the wall.  The bead of the thermocouple was flush with the surface on the calcium silcate board and was in contact with the board, as shown in Figure~\ref{Instrumented_Wall_Close_up_TC_HF_photo} For the center array,a second thermocouple was installed on the backside of the calcium silicate board, approximately 10~mm below the TC positioned on the front surface of the wall. The photograph in Figure~\ref{Instrumented_Wall_Detail_Rear_TC_photo} shows a rear thermocouple with the bead in contact with the rear surface of the wall and the relative position to the thermocouple on the front surface of the wall. In a manner similar to the source fire characterization experiments, the instrumented wall experiments were conducted under the NIST Large Fire Research Laboratory’s 3~m by 3~m oxygen depletion calorimeter.

@@

\section{Characterization of Gypsum WallBoard}

According to the United States Geological Survey, most gypsum in the United States is used to make wallboards for homes and other structures.  A typical home in the United States contains more than 7 metric tons of gypsum ref USGS website http://minerals.usgs.gov/minerals/pubs/commodity/gypsum/.  As defined by ASTM, gypsum wallboard is gypsum board designed for use on walls, ceilings, or partitions and that affords a surface suitable to recieve decoration  ref ATSM 1396. The ASTM standard further describes gypsum wallboard as having a noncombustible core, surfaced with paper bonded to the core. By weight, wallboard is composed of more than 85~\% gypsum, less than 10~\% cellulose, less than 3~\% starch and approximately 1~\% of both crystalline silica and fibrous glass.  Chemically gypsum can be described as calcium sulfate dihydrate, (CaSO$_{4}$  2H$_{2}O) .  By weight pure gypsum is approximately 79~\% calcium sulfate and 21~\% chemically bound water ref Lawson 1977.

@@

The first series of experiments were conducted with the natural gas burner.

 \begin{figure}[p]
 	\centering
 	\includegraphics[width=4.5in]{../Figures/IWGB_NG_TC_Surface_Center_Avg}\\
 	\includegraphics[width=4.5in]{../Figures/IWGB_NG_TC_Surface_Offset_Avg}\\
 	\includegraphics[width=4.5in]{../Figures/IWGB_NG_TC_Plume_Avg_Avg}\\
 	\caption[Averaged peak wall and plume temperatures for the natural gas burner]{Averaged peak wall and plume temperatures for the natural gas burner fires at several different positions from the instrumented gypsum board wall.  The positions ranged from 0.3 m away from the wall to physically touching the wall, 0.0 m.}
 	\label{IWGB_NG_Temp_FrontSet}
 \end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=4.5in]{../Figures/IWGB_NG_TC_Back_Center_Avg}\\
	\caption[Averaged peak temperatures on the backside of the gypsum board wall for the natural gas burner fires.]{Averaged peak temperatures along the vertical centerline of the backside of the gypsum board wall for the natural gas burner fires}
	\label{IWGB_NG_Temp_RearCenter}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=4.5in]{../Figures/IWGB_NG_HF_Center_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_NG_HF_Offset_Avg}\\
	\caption[Averaged peak heat flux on the gypsum board wall for the natural gas burner fires.]{Averaged peak heat flux on the gypsum board wall along the vertical centerline and edge of the natural gas burner fires. }
	\label{IWGB_NG_Temp_RearCenter}
\end{figure}

@@

Comparisons NG to Gas to PUF

\begin{figure}[p]
	\centering
	\includegraphics[width=4.5in]{../Figures/IWGB_NG_TC_Surface_Center_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_GAS_TC_Surface_Center_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_PUF_TC_Surface_Center_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
	\label{IWGB_Temp_Comp_FrontCenter}
\end{figure}

@@

\begin{figure}[p]
	\centering
	\includegraphics[width=4.5in]{../Figures/IWGB_NG_HF_Center_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_GAS_HF_Center_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_PUF_HF_Center_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
	\label{IWGB_HF_Comp_Center}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=4.5in]{../Figures/IWGB_NG_HF_Offset_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_GAS_HF_Offset_Avg}\\
	\includegraphics[width=4.5in]{../Figures/IWGB_PUF_HF_Offset_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
	\label{IWGB_HF_Comp_Edge}
\end{figure}

@@

\bibliography{../References/Madrzykowski_Thesis,../../../Bibliography/FDS_refs}
\end{document}
